---
title: "Tariffs"
format: pdf
editor: visual
---

# International economics

## Optimal tariffs and lobbying

```{r}
rm(list = ls())
library(tidyverse)
library(readxl)
# data
tariff <- read_excel("C:/Users/danie/Downloads/tariff.xlsx")
```

Use the spreadsheet ‘tariff.xls’ for the following questions:

(a) Suppose that the Home government maximizes the social welfare which is a
sum of consumer surplus, producer surplus and tariff revenue, CS+PS+TR



```{r}
# Variables I have are: `Home tariff`, HOME CS`, `HOME PS`, `HOME imports`

tariff <- tariff %>%
  mutate(
    SW = `HOME CS` + `HOME PS` + `Home tariff` * `HOME imports`
  )
```

Create a figure that plots Home social welfare on the vertical axis against the tariff level on the horizontal axis. Using this visual aide and data, find the optimal tariff

```{r}
ggplot(tariff, aes(x = `Home tariff`, y = SW)) +
  geom_line() +
  geom_point()+
  labs(title = "Social welfare vs. tariff level",
       x = "Tariff level",
       y = "Social welfare")+
  theme_minimal()+
  # Add a vertical line at the optimal tariff
  geom_vline(aes(xintercept = `Home tariff`[which.max(SW)]), linetype = "dashed", color = "red")+
  # Add a point at the optimal tariff
  geom_point(aes(x = `Home tariff`[which.max(SW)], y = max(SW)), color = "red", size = 3)+
  # fix the x-axis ticks
  scale_x_continuous(breaks = seq(0, 33.5, 1))

# Optimal tariff
tariff %>%
  filter(SW == max(SW))




```
Optimal tariff is when the social welfare is maximized. Social welfare is maximized at 1360.01 when the tariff is 9 cents.

Now you are told that the actual policy is the tariff level of 11 cents. What can
you learn from this comparison?

When the actual policy is the tariff level of 11 cents, the social welfare is 1359.99. This is very close to the optimal tariff level of 9 cents. This means that the actual policy is very close to the optimal policy.

(b) Next, examine how the ‘optimal’ tariff might change with a different
priority of the government. 

Suppose that the government is biased towards producers and puts an extra weight on PS. That is, the government maximizes CS+7PS+TR. What is the optimal tariff in this case?

```{r}

tariff <- tariff %>%
  mutate(
    SW2 = `HOME CS` + 7 * `HOME PS` + `Home tariff` * `HOME imports`
  )

ggplot(tariff, aes(x = `Home tariff`, y = SW2)) +
  geom_line() +
  geom_point()+
  labs(title = "Social welfare vs. tariff level",
       x = "Tariff level",
       y = "Social welfare")+
  theme_minimal()+
  # Add a vertical line at the optimal tariff
  geom_vline(aes(xintercept = `Home tariff`[which.max(SW2)]), linetype = "dashed", color = "red")+
  # Add a point at the optimal tariff
  geom_point(aes(x = `Home tariff`[which.max(SW2)], y = max(SW2)), color = "red", size = 3)+
  # fix the x-axis ticks
  scale_x_continuous(breaks = seq(0, 33.5, 1))

```


```{r}
# Optimal tariff
tariff %>%
  filter(SW2 == max(SW2))

```