---
format: 
  pdf:
    geometry:
      - top=20mm
      - left=20mm
      - heightrounded
    fontsize: 12pt
    documentclass: scrartcl
    papersize: a4
    toccolor: black
echo: false
warning: false
toc: true
lof: true
bibliography: documentobjects/texstuff/references.bib
csl: documentobjects/texstuff/apa.csl
header-includes:
  - \usepackage{wrapfig}
  - \usepackage{subcaption}
  - \usepackage{amsmath}
  - \usepackage{cancel}
  - \usepackage{hyperref}
  - \usepackage{tikz}
  - \usepackage{tabularx}
  - \usepackage{colortbl}
  - \usepackage{xcolor}
  - \renewcommand{\maketitle}{}
  - \definecolor{cornflowerblue}{RGB}{100,149,237}
  - \definecolor{darkgrey}{RGB}{220,220,220}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{} 
  - \fancyhead[L]{\rightmark}
  - \fancyhead[R]{\thepage}
  - \fancyfoot[C]{\thepage}
nocite: |
  @*
---

\newpage 

# Home assignment 1: Openness

The following questions are based on the Excel spreadsheet entitled “WorldBankData”. The data are from the World Bank. Countries with missing data have been omitted.

First, define the ‘**openness**’ of a country as the sum of its imports and exports divided by its GDP.

## (1) What has happened to the average level of openness in the world over the years in question?

```{r}
#Load the excel spreadsheet 
library(readxl) 
library(tidyverse)

data <- read_excel("WorldBankData.xlsx", skip = 1)

# Renaming alot of the data to fix the column names
data <- data %>% 
  rename(
    GDP_1971 = "GDP...2",
    GDP_1981 = "GDP...6",
    GDP_1991 = "GDP...10",
    GDP_2001 = "GDP...14",
    GDP_2011 = "GDP...18",
    GDP_2021 = "GDP...22",
    Exports_1971 = "Exports...3",
    Exports_1981 = "Exports...7",
    Exports_1991 = "Exports...11",
    Exports_2001 = "Exports...15",
    Exports_2011 = "Exports...19",
    Exports_2021 = "Exports...23",
    Imports_1971 = "Imports...4",
    Imports_1981 = "Imports...8",
    Imports_1991 = "Imports...12",
    Imports_2001 = "Imports...16",
    Imports_2011 = "Imports...20",
    Imports_2021 = "Imports...24",
    Population_1971 = "Population...5",
    Population_1981 = "Population...9",
    Population_1991 = "Population...13",
    Population_2001 = "Population...17",
    Population_2011 = "Population...21",
    Population_2021 = "Population...25"
  )
  

# Calculating the "openness" of the economy of each country for each tenth year
data <- data %>% 
  mutate(
    Openness_1971 = (Exports_1971 + Imports_1971) / (GDP_1971),
    Openness_1981 = (Exports_1981 + Imports_1981) / (GDP_1981),
    Openness_1991 = (Exports_1991 + Imports_1991) / (GDP_1991),
    Openness_2001 = (Exports_2001 + Imports_2001) / (GDP_2001),
    Openness_2011 = (Exports_2011 + Imports_2011) / (GDP_2011),
    Openness_2021 = (Exports_2021 + Imports_2021) / (GDP_2021)
  )

# Calculating an average oppennes for years rounded with two decimals
data <- data %>% 
  mutate(Average_1971 = round(mean(Openness_1971),2),
         Average_1981 = round(mean(Openness_1981),2),
         Average_1991 = round(mean(Openness_1991),2),
         Average_2001 = round(mean(Openness_2001),2),
         Average_2011 = round(mean(Openness_2011),2),
         Average_2021 = round(mean(Openness_2021),2)
  )

# Calculate the change of opennes from 2011 to 2022
data <- data %>% 
  mutate(
    Change_2011_2021 = Openness_2021 - Openness_2011
  )
```


```{r}
# Plotting average openness over the years
fig_openness<-data %>%
  select(Average_1971, Average_1981, Average_1991, Average_2001, Average_2011, Average_2021) %>%
  pivot_longer(everything(), names_to = "Year", values_to = "Openness") %>%
  ggplot(aes(x = Year, y = Openness, group = 1)) +
  geom_line(color = "cornflowerblue", linewidth = 1.2) +
  geom_point(color = "cornflowerblue", size = 3) +
  labs(title = "",
       x = "Year",
       y = "Openness") +
  # annotate the plot with the values of the years
  geom_text(aes(label = Openness), nudge_y = 0.03, check_overlap = TRUE, size = 7) +
    theme_bw()+
  # make text for x and y points bigger
  theme(axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))

#ggsave("documentobjects/figures/openness.png", plot = fig_openness, width = 10, height = 6, dpi = 300)
fig_openness
```


```{=tex}
\begin{figure}[H]
\centering
\includegraphics{documentobjects/figures/openness.png}
\caption{Average openness over the years}
\end{figure}
```

The average openness over the years has increased from 1971 to 2021 but has had a small decrease the last 10 years after a peak of 0.87 in 2011.


## (2) How many countries experienced an increase in openness between 2011 and 2021?

```{r}
# Count the countries with an increase in opennes from 2011 to 2021
data %>%
  filter(Change_2011_2021 > 0) %>%
  count()
```

Thirty six countries experienced an increase in openness between 2011 and 2021.

\newpage

```{r}
# Calculate gdp per capita
data <- data %>% 
  mutate(
    GDP_per_capita_1971 = GDP_1971 / Population_1971,
    GDP_per_capita_1981 = GDP_1981 / Population_1981,
    GDP_per_capita_1991 = GDP_1991 / Population_1991,
    GDP_per_capita_2001 = GDP_2001 / Population_2001,
    GDP_per_capita_2011 = GDP_2011 / Population_2011,
    GDP_per_capita_2021 = GDP_2021 / Population_2021
  )
```

## (3) Compare the average openness of the 20 largest countries (measured in terms of population) and the 20 smallest countries in 2011. Which is more open?


```{r}
# Calculate the average openness of the 20 largest and 20 smallest countries in 2011
sjekk_3 <-data %>%
  arrange(desc(Population_2011)) %>%
  slice_head(n = 20) %>%
  summarise(Average_openness_largest = mean(Openness_2011))

round(sjekk_3,2)
```

```{r}
data %>%
  arrange(Population_2011) %>%
  slice_head(n = 20) %>%
  summarise(Average_openness_smallest = mean(Openness_2011))
```
The smallest countries are more open than the largest countries in 2011.


## (4) Now compare the average openness of the 20 richest countries (measured in terms of per capita GDP) and the 20 poorest countries in 2011. Which is more open?

```{r}
# Calculate the average openness of the 20 richest and 20 poorest countries in 2011
sjekk_1 <- data %>%
  arrange(desc(GDP_per_capita_2011)) %>%
  slice_head(n = 20) %>%
  summarise(Average_openness_richest = mean(Openness_2011))

round(sjekk_1,2)
```

```{r}
sjekk <-data %>%
  arrange(GDP_per_capita_2011) %>%
  slice_head(n = 20) %>%
  summarise(Average_openness_poorest = mean(Openness_2011))

round(sjekk,2)
```
The richest countries are more open than the poorest countries in 2011.

## (5) Now, the most important part of your empirical analysis! Based on your results, what kind of countries tend to be more open and what kind tend to be less open.

The results show that the smallest and richest countries tend to be more open than the largest and poorest countries.

\newpage


