---
format: 
  pdf:
    geometry:
      - top=20mm
      - left=20mm
      - heightrounded
    fontsize: 12pt
    documentclass: scrartcl
    papersize: a4
    toccolor: black
echo: false
warning: false
toc: true
lof: true
bibliography: documentobjects/texstuff/references.bib
csl: documentobjects/texstuff/apa.csl
header-includes:
  - \usepackage{wrapfig}
  - \usepackage{subcaption}
  - \usepackage{amsmath}
  - \usepackage{cancel}
  - \usepackage{hyperref}
  - \usepackage{tikz}
  - \usepackage{tabularx}
  - \usepackage{colortbl}
  - \usepackage{xcolor}
  - \renewcommand{\maketitle}{}
  - \definecolor{cornflowerblue}{RGB}{100,149,237}
  - \definecolor{darkgrey}{RGB}{220,220,220}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{} 
  - \fancyhead[L]{\rightmark}
  - \fancyhead[R]{\thepage}
  - \fancyfoot[C]{\thepage}
nocite: |
  @*
---

\newpage 

# Home assignment 5: The Melitz effect


The following questions are based on the Excel spreadsheet entitled “heterogeousfirms.xls”. This provides data for a hypothetical monopolistically competitive market with heterogeneous firms. 

Each firm is numbered from 1 to 100, and has its marginal product of labor φ marked. The common value of the fixed labor requirement, f, is marked at the top of the spreadsheet. For each firm, an assumed value for the firm’s initial quantity produced is marked as well; assume that this has been derived by setting each firm’s marginal cost equal to its marginal revenue. Fixed cost for production f = 50. And fixed cost for exporting g = 25.

Note that firms with higher marginal products of labor are assumed to produce more output.

```{r}
#Load the excel spreadsheet 
rm(list=ls())
library(readxl) 
library(tidyverse)

url <- "https://raw.githubusercontent.com/DanielFabioG/data/main/heterogeous_firms.xls"
temp <- tempfile(fileext = ".xls")

# Download the file to a temporary location
download.file(url, temp, mode = "wb")

# Read the Excel file and cut first 3 rows
data <- read_excel(temp, skip = 4)
rm(temp, url)
```

(a) Compute each firm’s employment of labor under autarky.

```{r}
f <- 50
g <- 25

# Calculate the employment of labor under autarky
data <- data %>%
  mutate(Labor_autarky = f + `Quantity produced under autarky.` / `Marginal product of labor, Φ.`)

# Show the first 10 rows of the data
head(data, 10)
```
(b) Use this information to compute the industry’s labor productivity (total output per
worker).

```{r}
# Calculate the industry's labor productivity
industry_labor_productivity <- sum(data$`Quantity produced under autarky.`) / sum(data$Labor_autarky)

# Round the result to 2 decimals
round(industry_labor_productivity, 2)
```
(c) Now, suppose that the industry is opened to trade, and in accordance with the Melitz effect, the least efficient 15% of the firms drop out. Further, suppose that firm #54 and all of the firms more efficient than firm #54 export, while the remainder of the surviving firms produce only for the domestic market. Suppose that exporting firms increase their output by 10% compared to autarky, while non-exporters reduce their output by 10% compared to autarky. Now, redo your calculations in parts (a) and (b). Interpret your results. In particular, what happens to industry productivity and why?

```{r}
# Lets make a new column where the 15 % least efficient firms are marked as "dropouts"
# the middle will be marked as "domestic" and the most efficient firms as "exporters"
data <- data %>%
  mutate(Export = ifelse(row_number() <= 15, "Dropouts",
                          ifelse(row_number() <= 54, "Domestic", "Exporters")))

```































```{r}
# Plotting average openness over the years
fig_openness<-data %>%
  select(Average_1971, Average_1981, Average_1991, Average_2001, Average_2011, Average_2021) %>%
  pivot_longer(everything(), names_to = "Year", values_to = "Openness") %>%
  ggplot(aes(x = Year, y = Openness, group = 1)) +
  geom_line(color = "cornflowerblue", linewidth = 1.2) +
  geom_point(color = "cornflowerblue", size = 3) +
  labs(title = "",
       x = "Year",
       y = "Openness") +
  # annotate the plot with the values of the years
  geom_text(aes(label = Openness), nudge_y = 0.03, check_overlap = TRUE, size = 7) +
    theme_bw()+
  # make text for x and y points bigger
  theme(axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))

#ggsave("documentobjects/figures/openness.png", plot = fig_openness, width = 10, height = 6, dpi = 300)
fig_openness
```


```{=tex}
\begin{figure}[H]
\centering
\includegraphics{documentobjects/figures/openness.png}
\caption{Average openness over the years}
\end{figure}
```

The average openness over the years has increased from 1971 to 2021 but has had a small decrease the last 10 years after a peak of 0.87 in 2011.


## (2) How many countries experienced an increase in openness between 2011 and 2021?

```{r}
# Count the countries with an increase in opennes from 2011 to 2021
data %>%
  filter(Change_2011_2021 > 0) %>%
  count()
```

Thirty six countries experienced an increase in openness between 2011 and 2021.

\newpage

```{r}
# Calculate gdp per capita
data <- data %>% 
  mutate(
    GDP_per_capita_1971 = GDP_1971 / Population_1971,
    GDP_per_capita_1981 = GDP_1981 / Population_1981,
    GDP_per_capita_1991 = GDP_1991 / Population_1991,
    GDP_per_capita_2001 = GDP_2001 / Population_2001,
    GDP_per_capita_2011 = GDP_2011 / Population_2011,
    GDP_per_capita_2021 = GDP_2021 / Population_2021
  )
```

## (3) Compare the average openness of the 20 largest countries (measured in terms of population) and the 20 smallest countries in 2011. Which is more open?


```{r}
# Calculate the average openness of the 20 largest and 20 smallest countries in 2011
sjekk_3 <-data %>%
  arrange(desc(Population_2011)) %>%
  slice_head(n = 20) %>%
  summarise(Average_openness_largest = mean(Openness_2011))

round(sjekk_3,2)
```

```{r}
data %>%
  arrange(Population_2011) %>%
  slice_head(n = 20) %>%
  summarise(Average_openness_smallest = mean(Openness_2011))
```
The smallest countries are more open than the largest countries in 2011.


## (4) Now compare the average openness of the 20 richest countries (measured in terms of per capita GDP) and the 20 poorest countries in 2011. Which is more open?

```{r}
# Calculate the average openness of the 20 richest and 20 poorest countries in 2011
sjekk_1 <- data %>%
  arrange(desc(GDP_per_capita_2011)) %>%
  slice_head(n = 20) %>%
  summarise(Average_openness_richest = mean(Openness_2011))

round(sjekk_1,2)
```

```{r}
sjekk <-data %>%
  arrange(GDP_per_capita_2011) %>%
  slice_head(n = 20) %>%
  summarise(Average_openness_poorest = mean(Openness_2011))

round(sjekk,2)
```
The richest countries are more open than the poorest countries in 2011.

## (5) Now, the most important part of your empirical analysis! Based on your results, what kind of countries tend to be more open and what kind tend to be less open.

The results show that the smallest and richest countries tend to be more open than the largest and poorest countries.

\newpage


